---
title: "How to Analyze Memory Scores in Psychology"
layout: post
output:
  md_document:
    toc: yes
    variant: markdown_github
  html_document:
    df_print: paged
    toc: yes
author: "Dominique Makowski"
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---


```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
```


# The Problem

# The data

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
library(tidyverse)  # Load a collection of useful packages 
library(psycho)  # Load the psycho package

df <- psycho::emotion %>% 
  select(Participant_ID, Participant_Age, Participant_Sex, Emotion_Condition, Autobiographical_Link, Recall)

summary(df)
```

We started by loading an available dataset from the psycho package, selecting variables that interest us, and storing the dataframe into an object called `df`. Then, to have a glimpse at the dataframe, we ran the `summary` function on it, that shows us the different variables, their type and some info about them.



# Intra-individual Analysis

## The Effect of Emotion on Free Recall

### The Primitive Way


```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
df_aov <- df %>% 
  mutate(Recall = ifelse(Recall == TRUE, 1, 0)) %>% 
  group_by(Participant_ID, Emotion_Condition) %>% 
  summarise(Recall_Ratio = sum(Recall) / n())

head(df_aov)  # We can see the 10 first rows of a dataframe by running the "head()" function
```

### ANOVA

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
aov_result <- aov(Recall_Ratio ~ Emotion_Condition + Error(Participant_ID), data=df_aov)
analyze(aov_result)
```


### Mixed Model

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
library(lmerTest)

fit <- lmer(Recall_Ratio ~ Emotion_Condition + (1|Participant_ID), data=df_aov)

aov_result <- anova(fit)
analyze(aov_result)
```

As you can see, the results are close. Nevertheless, it's better to analyze the full model.

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
fit <- lmer(Recall_Ratio ~ Emotion_Condition + (1|Participant_ID), data=df_aov)
analyze(fit)
```


**And yet, what we just is wrong! Indeed, we treated ratios (that have 0-1 bounds) as regular **

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
Zdf_aov <- standardize(ungroup(df_aov))
fit <- lmer(Recall_Ratio ~ Emotion_Condition + (1|Participant_ID), data=Zdf_aov)
analyze(fit)

fit <- rstanarm::stan_glmer(Recall_Ratio ~ Emotion_Condition + (1|Participant_ID), data=Zdf_aov)
analyze(fit)



dummies <- scale(contrasts(df$Species)[df$Species,])
fit <- lm(Sepal.Length ~ dummies, data = df)
summary(fit)
round(coef(fit), 2)
round(MuMIn::std.coef(fit, partial.sd = FALSE), 2)


library(tidyverse)
library(MuMIn) # For stds coefs

df <- iris %>% 
  select(Sepal.Length, Sepal.Width) %>% 
  scale() %>% 
  as.data.frame() %>% 
  mutate(Species = iris$Species)

fit <- lm(Sepal.Length ~ Sepal.Width, data=df)
round(coef(fit), 2)
round(MuMIn::std.coef(fit, partial.sd = TRUE), 2)

fit <- lm(Sepal.Length ~ Species, data=df)
round(coef(fit), 2)
round(MuMIn::std.coef(fit, partial.sd = TRUE), 2)




library(rstanarm)

std_posteriors <- function(fit){
      X <- model.matrix(fit)
    sd_X <- apply(X, MARGIN = 2, FUN = sd)[-1]
    sd_Y <- apply(posterior_predict(fit), MARGIN = 1, FUN = sd)
    beta <- as.matrix(fit)[, 2:ncol(X), drop = FALSE]
    posteriors_std <- sweep(
      sweep(beta, MARGIN = 2, STATS = sd_X, FUN = `*`),
      MARGIN = 1, STATS = sd_Y, FUN = `/`
    )
    return(posteriors_std)
}

df <- iris %>% 
  select(Sepal.Length, Sepal.Width) %>% 
  scale() %>% 
  as.data.frame() %>% 
  mutate(Species = iris$Species)
  
  

fit <- rstanarm::stan_glm(Sepal.Length ~ Species, data=df)
post <- data.frame(fit)$Speciesversicolor
std_post <- as.data.frame(std_posteriors(fit))$Speciesversicolor

paste("Median = ", round(median(post), 2), ", MAD = ", round(mad(post), 2))
paste("Std. Median = ", round(median(std_post), 2), ", Std. MAD = ", round(mad(std_post), 2))

```

# Inter-individual Analysis

## Signal Detection Theory

See THIS blogpost.





# Credits

The psycho package helped you? Don't forget to cite the various packages you used :)

You can cite `psycho` as follows:

- Makowski, (2018). *The psycho Package: An Efficient and Publishing-Oriented Workflow for Psychological Science*. Journal of Open Source Software, 3(22), 470. https://doi.org/10.21105/joss.00470


# Previous blogposts

- [Formatted Correlation with Effect Size](https://neuropsychology.github.io/psycho.R/2018/06/28/analyze_correlation.html)
- [Extracting a Reference Grid of your Data for Machine Learning Models Visualization](https://neuropsychology.github.io/psycho.R/2018/06/25/refdata.html)
- [Copy/paste t-tests Directly to Manuscripts](https://neuropsychology.github.io/psycho.R/2018/06/19/analyze_ttest.html)
- [Easy APA Formatted Bayesian Correlation](https://neuropsychology.github.io/psycho.R/2018/06/11/bayesian_correlation.html)
- [Fancy Plot (with Posterior Samples) for Bayesian Regressions](https://neuropsychology.github.io/psycho.R/2018/06/03/plot_bayesian_model.html)
- [How Many Factors to Retain in Factor Analysis](https://neuropsychology.github.io/psycho.R/2018/05/24/n_factors.html)
- [Beautiful and Powerful Correlation Tables](https://neuropsychology.github.io/psycho.R/2018/05/20/correlation.html)
- [Format and Interpret Linear Mixed Models](https://neuropsychology.github.io/psycho.R/2018/05/10/interpret_mixed_models.html)
- [How to do Repeated Measures ANOVAs](https://neuropsychology.github.io/psycho.R/2018/05/01/repeated_measure_anovas.html)
- [Standardize (Z-score) a dataframe](https://neuropsychology.github.io/psycho.R/2018/03/29/standardize.html)
- [Compute Signal Detection Theory Indices](https://neuropsychology.github.io/psycho.R/2018/03/29/SDT.html)
- [Installing R, R Studio and psycho](https://neuropsychology.github.io/psycho.R/2018/03/21/installingR.html)
