---
title: "How to Analyze Memory Scores in Psychology"
layout: post
output:
  md_document:
    toc: yes
    variant: markdown_github
  html_document:
    df_print: paged
    toc: yes
author: "Dominique Makowski"
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---


```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
```


# The Problem

# The data

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
library(tidyverse)  # Load a collection of useful packages 
library(psycho)  # Load the psycho package

df <- psycho::emotion %>% 
  select(Participant_ID, Participant_Age, Participant_Sex, Emotion_Condition, Autobiographical_Link, Recall)

summary(df)
```

We started by loading an available dataset from the psycho package, selecting variables that interest us, and storing the dataframe into an object called `df`. Then, to have a glimpse at the dataframe, we ran the `summary` function on it, that shows us the different variables, their type and some info about them.



# Intra-individual Analysis

## The Effect of Emotion on Free Recall

### The Primitive Way


```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
df_aov <- df %>% 
  mutate(Recall = ifelse(Recall == TRUE, 1, 0)) %>% 
  group_by(Participant_ID, Emotion_Condition) %>% 
  summarise(Recall_Ratio = sum(Recall) / n())

head(df_aov)  # We can see the 10 first rows of a dataframe by running the "head()" function
```

### ANOVA

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
aov_result <- aov(Recall_Ratio ~ Emotion_Condition + Error(Participant_ID), data=df_aov)
analyze(aov_result)
```


### Mixed Model

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
library(lmerTest)

fit <- lmerTest::lmer(Recall_Ratio ~ Emotion_Condition + (1|Participant_ID), data=df_aov)

aov_result <- anova(fit)
analyze(aov_result)
```

You can note that in this version, no effect sizes are returned. This is because they are bad and make no really sense for mixed (repeated measures) designs.

As you can see, the results are close. Nevertheless, it's better to analyze the full model.

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
library(lme4)

fit <- lmerTest::lmer(Recall_Ratio ~ Emotion_Condition + (1|Participant_ID), data=df_aov)
analyze(fit)
```


**And yet, what we just is wrong! Indeed, we treated ratios (that have 0-1 bounds) as regular **

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
fit <- lme4::glmer(Recall ~ Emotion_Condition + (1|Participant_ID), data=df, family="binomial")
analyze(fit)
```

### Bayesian
```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
fit <- rstanarm::stan_glmer(Recall ~ Emotion_Condition + (1|Participant_ID), data=df, family="binomial", chains=1)
analyze(fit, effsize=TRUE)

fit <- rstanarm::stan_glmer(Recall ~ Autobiographical_Link + (1|Participant_ID), data=df, family="binomial", chains=1)
analyze(fit, effsize=TRUE)
```


# Inter-individual Analysis

## Signal Detection Theory

See THIS blogpost.





# Credits

The psycho package helped you? Don't forget to cite the various packages you used :)

You can cite `psycho` as follows:

- Makowski, (2018). *The psycho Package: An Efficient and Publishing-Oriented Workflow for Psychological Science*. Journal of Open Source Software, 3(22), 470. https://doi.org/10.21105/joss.00470


# Previous blogposts

- [Formatted Correlation with Effect Size](https://neuropsychology.github.io/psycho.R/2018/06/28/analyze_correlation.html)
- [Extracting a Reference Grid of your Data for Machine Learning Models Visualization](https://neuropsychology.github.io/psycho.R/2018/06/25/refdata.html)
- [Copy/paste t-tests Directly to Manuscripts](https://neuropsychology.github.io/psycho.R/2018/06/19/analyze_ttest.html)
- [Easy APA Formatted Bayesian Correlation](https://neuropsychology.github.io/psycho.R/2018/06/11/bayesian_correlation.html)
- [Fancy Plot (with Posterior Samples) for Bayesian Regressions](https://neuropsychology.github.io/psycho.R/2018/06/03/plot_bayesian_model.html)
- [How Many Factors to Retain in Factor Analysis](https://neuropsychology.github.io/psycho.R/2018/05/24/n_factors.html)
- [Beautiful and Powerful Correlation Tables](https://neuropsychology.github.io/psycho.R/2018/05/20/correlation.html)
- [Format and Interpret Linear Mixed Models](https://neuropsychology.github.io/psycho.R/2018/05/10/interpret_mixed_models.html)
- [How to do Repeated Measures ANOVAs](https://neuropsychology.github.io/psycho.R/2018/05/01/repeated_measure_anovas.html)
- [Standardize (Z-score) a dataframe](https://neuropsychology.github.io/psycho.R/2018/03/29/standardize.html)
- [Compute Signal Detection Theory Indices](https://neuropsychology.github.io/psycho.R/2018/03/29/SDT.html)
- [Installing R, R Studio and psycho](https://neuropsychology.github.io/psycho.R/2018/03/21/installingR.html)
