---
title: "APA Formatted Correlation"
layout: post
output:
  html_document:
    df_print: paged
    toc: yes
  md_document:
    toc: yes
    variant: markdown_github
author: "Dominique Makowski"
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---


```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
```

One of the most time-consuming part of data analysis in psychology is the copy-pasting of specific values of some R outputs to a manuscript or a report. This task is frustrating, prone to errors, and increase de variability of statistical reporting, while standardization of what and how to report is crucial for reproducibility and clarity. The `psycho` package was designed specifically to ease this part, at first for complex Bayesian mixed models, but is now also compatible with basic methods, such as **correlation**.

# Do a correlation
```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
df <- iris
cor_results <- cor.test(df$Sepal.Length, df$Petal.Length)
```


# APA formatted output

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
# devtools::install_github("neuropsychology/psycho.R")  # Install the latest psycho version

# Load packages
library(psycho)
psycho::analyze(cor_results)
```


The formatted output includes **direction**, **effect size** (interpreted by default with **Cohen's (1988)** rules) and **confidence intervals**. Now you can just copy and paste this line into your report and focus on what's important.

# Dataframe of Values

It is also possible to have all the values stored in a dataframe by running a **summary** on the analyzed object.

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='hide', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
results <- analyze(cor_results)
summary(results)
```
```{r, fig.width=7, fig.height=4.5, eval = TRUE, echo=FALSE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
knitr::kable(summary(results), digits=2)
```

# Bayesian Correlation

Nevertheless, we recommand doing a **Bayesian correlation**, which is even [easier and quicker to do](blogpostlink)!



# Credits

This package helped you? Don't forget to cite the various packages you used :)

You can cite `psycho` as follows:

- Makowski, (2018). *The psycho Package: An Efficient and Publishing-Oriented Workflow for Psychological Science*. Journal of Open Source Software, 3(22), 470. https://doi.org/10.21105/joss.00470
