---
title: "GLMs: Nested Effects or Interaction?"
layout: post
output:
  md_document:
    toc: yes
    variant: markdown_github
  html_document:
    df_print: paged
    toc: yes
author: "Dominique Makowski"
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---


```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
```

When I started doing statistics, after a bachelor in psychology, there was, to me, no such things as nested effects. You tested a model (often directly the ANOVA) with the interaction or without. 

Let's take interest in the relationship between **life satisfaction**, **sex** and **concealing** (the tendency to suppress or hide one's emotions)

# The Data

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
library(tidyverse)
library(psycho)

df <- psycho::affective %>% 
  select(Sex, Life_Satisfaction, Adjusting)

summary(df)
```

# Primitve Way
```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
anova_results <- aov(Life_Satisfaction ~ Sex * Concealing, data=df)
analyze(anova_results)
```

If we wanted to go deeper, we have to do something like post-hocs tests and stuff, which are messy and not the correct way to proceed. Instead, we will investigate the underlying linear regression model.


# Regression 
```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
fit <- lm(Life_Satisfaction ~ Sex * Concealing, data=df)
```

First, note that running an anova on the linear model produces EXACTLY the same results:
```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
analyze(anova(fit)) 
```

 But we can to have a look at the model itself, which is richer than the ANOVA.

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
analyze(fit) 
```

Again, three lines of "effect". One could think these a similar to the ones of ANOVA, with the main effect of Sex, the main effect of concealing and the interaction. But they are not exactly that! they refer to specific parameters (or "effects") of the model. 

# Visualize the data 


```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
df %>% 
  ggplot(aes(x=Concealing, y=Life_Satisfaction, colour=Sex)) +
  geom_point() +
  geom_smooth(method="lm")
```

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
df %>% 
  ggplot(aes(x=Concealing, y=Life_Satisfaction, colour=Sex)) +
  geom_jitter(width = 0.1, height = 0.1, alpha=0.5) +
  geom_smooth(method="lm")
```

# Visualize the model

This method is alright for simple models, but for more complex kind of stuffs wouldn't work, as your model infered things that cannot be directly represented with the data on a 2D panel. Nevertheless, one can directly plot the model's prediction.

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
refdata <- df %>% 
  select(Sex, Concealing) %>% 
  refdata()

get_predicted(fit, newdata=refdata) %>% 
  ggplot(aes(x=Concealing, y=Life_Satisfaction_Predicted)) +
  geom_line(aes(colour=Sex)) +
  geom_ribbon(aes(fill=Sex,
                  ymin=Life_Satisfaction_CI_2.5,
                  ymax=Life_Satisfaction_CI_97.5),
              alpha=0.25)
```

This is what our model predicted. 


# Effects

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
refdata <- df %>% 
  select(Sex, Concealing) %>% 
  refdata()

get_predicted(fit, newdata=refdata) %>% 
  ggplot(aes(x=Concealing, y=Life_Satisfaction_Predicted)) +
  geom_line(aes(colour=Sex)) +
  geom_ribbon(aes(fill=Sex,
                  ymin=Life_Satisfaction_CI_2.5,
                  ymax=Life_Satisfaction_CI_97.5),
              alpha=0.25) + 
  geom_segment(aes(x = 0, y = 5.19, xend =0.2, yend = 5.19), size = 3, color="#f44336") +
  geom_segment(aes(x = 0, y = 5.1, xend =0, yend = 4.6), size = 3, arrow = arrow(length = unit(0.3, "inches")), color="#2196F3") +
  geom_segment(aes(x = 4, y = 4.8, xend =6, yend = 4.6), size = 3, arrow = arrow(length = unit(0.3, "inches")), color="#4CAF50") +
  geom_curve(aes(x = 6, y = 4.65, xend =6, yend = 5), size = 3, arrow = arrow(length = unit(0.3, "inches")), color="#FFC107")
```

These four parameters are sufficient to fully describe the model. And if they are "significantly" different from 0.

# Changing Effects

Nevertheless, sometimes, these effects are not exactly what we look for. For example, I could be specifically interested in the value and significance of the blue slope (the relationship between concealing and life satisfaction in men). With this model, I only know that this slope is indeed different than the women's one. But there could be an absence of significance in men, and a strongly significant link in women. Or a significantly opposite link in men? How can we know this? How can be change the effects displayed?

Wait for the next post about changing reference levels and nested effects!


# Contribute

Of course, these reporting standards should change, depending on new expert recommandations or official guidelines. **The goal of this package is to flexibly adaptive to new changes and good practices evolution**. Therefore, if you have any advices, opinions or such, we encourage you to either let us know by opening an [issue](https://github.com/neuropsychology/psycho.R/issues), or even better, try to implement them yourself by [contributing](https://github.com/neuropsychology/psycho.R/blob/master/.github/CONTRIBUTING.md) to the code. 


# Credits

This package helped you? Don't forget to cite the various packages you used :)

You can cite `psycho` as follows:

- Makowski, (2018). *The psycho Package: An Efficient and Publishing-Oriented Workflow for Psychological Science*. Journal of Open Source Software, 3(22), 470. https://doi.org/10.21105/joss.00470


# Previous blogposts

- [APA Formatted Bayesian Correlation](https://neuropsychology.github.io/psycho.R/2018/06/11/bayesian_correlation.html)
- [Fancy Plot (with Posterior Samples) for Bayesian Regressions](https://neuropsychology.github.io/psycho.R/2018/06/03/plot_bayesian_model.html)
- [How Many Factors to Retain in Factor Analysis](https://neuropsychology.github.io/psycho.R/2018/05/24/n_factors.html)
- [Beautiful and Powerful Correlation Tables](https://neuropsychology.github.io/psycho.R/2018/05/20/correlation.html)
- [Format and Interpret Linear Mixed Models](https://neuropsychology.github.io/psycho.R/2018/05/10/interpret_mixed_models.html)
- [How to do Repeated Measures ANOVAs](https://neuropsychology.github.io/psycho.R/2018/05/01/repeated_measure_anovas.html)
- [Standardize (Z-score) a dataframe](https://neuropsychology.github.io/psycho.R/2018/03/29/standardize.html)
- [Compute Signal Detection Theory Indices](https://neuropsychology.github.io/psycho.R/2018/03/29/SDT.html)
