---
title: "Very Easy Bayesian Correlation"
layout: post
output:
  html_document:
    df_print: paged
    toc: yes
  md_document:
    toc: yes
    variant: markdown_github
author: "Dominique Makowski"
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---


```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
```

**The Bayesian framework is the right way to go for psychological science.** To facilitate its use for newcommers, we implemented the `bayesian_cor.test` function in the [`psycho` package](https://github.com/neuropsychology/psycho.R), a **user-friendly wrapper** for the `correlationBF` function of the great [`BayesFactor`](https://richarddmorey.github.io/BayesFactor/) package by Richard D. Morey.

# Traditional Correlation

Let's perform a traditional correlation.

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
# devtools::install_github("neuropsychology/psycho.R")  # Install the latest psycho version

# Load packages
library(tidyverse)
library(psycho)

# Import data
df <- psycho::affective

cor.test(df$Concealing, df$Tolerating)
```

# Bayesian APA formatted Correlation

And now, see how quickly we can do a Bayesian correlation: 

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
bayes_cor.test(df$Concealing, df$Tolerating)
```

You can, now, simply copy/paste this output to your manuscript! It includes the Bayes Factor, the median (a good point-estimates, close to the *r* estimated in frequentist correlation), MAD (robust equivalent of SD for median) and *credible* interval (CI) of the posterior distribution, as well as effect size interpretation.

# Indices

We can have access to more indices with the `summary`:

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='hide', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
results <- bayes_cor.test(df$Concealing, df$Tolerating)
summary(results)
```
```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='hide', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
knitr::kable(summary(results), digits=2)
```

Those indices include the ROPE decision criterion (see REFERENCE) as well as the Maximum Probability of Effect (MPE, which equals the sum of probabilities of all effect sizes).




# Credits

This package helped you? Don't forget to cite the various packages you used :)

You can cite `psycho` as follows:

- Makowski, (2018). *The psycho Package: an Efficient and Publishing-Oriented Workflow for Psychological Science*. Journal of Open Source Software, 3(22), 470. https://doi.org/10.21105/joss.00470
