---
title: "Very Easy Bayesian Correlation"
layout: post
output:
  html_document:
    df_print: paged
    toc: yes
  md_document:
    toc: yes
    variant: markdown_github
author: "Dominique Makowski"
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---


```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
```

The Bayesian framework is the right way to go for psychological science. with the `bayesian_cor.test` function, from the [`psycho` package](https://github.com/neuropsychology/psycho.R), a user-friendly wrapper for the `correlationBF` function of the great [`BayesFactor`](https://richarddmorey.github.io/BayesFactor/) package by Richard D. Morey.

# Formatted Output

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
# devtools::install_github("neuropsychology/psycho.R")  # Install the last psycho version if needed

# Load packages
library(tidyverse)
library(psycho)

# Import data
df <- psycho::affective

results <- psycho::bayes_cor.test(df$Concealing, df$Tolerating)
print(results)
```

# Indices

We can have access to more indices with the `summary`:

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='hide', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
summary(results)
```
```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='hide', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
knitr::kable(summary(results))
```

Those indices include the ROPE decision criterion (see REFERENCE) as well as the Maximum Probability of Effect (MPE, which equals the sum of probabilities of all effect sizes).


# Posterior Distribution

Those indices are computed from the posterior distribution of the correlation values.



# Credits

This package helped you? Don't forget to cite the various packages you used :)

You can cite `psycho` as follows:

- Makowski, (2018). *The psycho Package: an Efficient and Publishing-Oriented Workflow for Psychological Science*. Journal of Open Source Software, 3(22), 470. https://doi.org/10.21105/joss.00470
