---
title: "Formatted and Flexible Correlation Tables"
layout: post
output:
  html_document:
    df_print: paged
    toc: yes
  md_document:
    toc: yes
    variant: markdown_github
author: "Dominique Makowski"
date: "`r Sys.Date()`"
tags:
- r
- psychology
- neuroscience
- neuropsychology
---


```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
```

# Another *correlation* function?!

Yes, the `correlation` function from the [`psycho`](https://github.com/neuropsychology/psycho.R) package.

```{r fig.align='center', message=FALSE, warning=FALSE, comment=NA, results='hide'}
library(psycho)
library(tidyverse)

cor <- psycho::affective %>% 
  correlation()
```

This function automatically select numeric variables and run a correlation analysis. It returns a `psychobject`.


# A table

We can then extract a **formatted table** that can be saved and pasted into reports and manuscripts.

```{r fig.align='center', message=FALSE, warning=FALSE, comment=NA, results='hide'}
summary(cor)
# write.csv(summary(cor), "myformattedcortable.csv")
```
```{r echo=FALSE, fig.align='center', message=FALSE, warning=FALSE, comment=NA, results='markup'}
knitr::kable(summary(cor))
```

# A Plot

It integrates a **plot** done with [`ggcorplot`](https://github.com/kassambara/ggcorrplot).

```{r, eval = TRUE, results='markup', fig.align='center'}
plot(cor)
```


# A print

It also includes a **pairwise correlation printing** method.

```{r fig.align='center', message=FALSE, warning=FALSE, comment=NA, results='markup'}
print(cor)
```

# Options

You can also cutomize the **type** (pearson, spearman or kendall), the **p value correction method** (holm (default), bonferroni, fdr, none...) and run **partial, semi-partial or glasso** correlations.


```{r fig.align='center', message=FALSE, warning=FALSE, comment=NA, results='hide'}
psycho::affective %>% 
  correlation(method = "pearson", adjust="bonferroni", type="partial") %>% 
  summary()
```
```{r echo=FALSE, fig.align='center', message=FALSE, warning=FALSE, comment=NA, results='markup'}
psycho::affective %>% 
  correlation(method = "pearson", adjust="bonferroni", type="partial") %>% 
  summary() %>% 
  knitr::kable()
```

# Fun with *p*-hacking

In order to **prevent people for running many uncorrected correlation tests** (promoting *p*-hacking and result-fishing), we included the `i_am_cheating` parameter. If FALSE (default), the function will help you finding interesting results!

```{r, eval = TRUE, results='hide', fig.align='center'}
df_with_11_vars <- data.frame(replicate(11, rnorm(1000)))
cor <- correlation(df_with_11_vars, adjust="none") 
summary(cor)
```
```{r echo=FALSE, fig.align='center', message=FALSE, warning=FALSE, comment=NA, results='markup'}
knitr::kable(summary(cor)[,1:11])
```


**As we can see, Schopenhauer's Optimism is strongly related to many variables!!!**


# Credits

This package was useful? You can cite [`psycho`](https://github.com/neuropsychology/psycho.R) as follows:

- Makowski, (2018). *The psycho Package: an Efficient and Publishing-Oriented Workflow for Psychological Science*. Journal of Open Source Software, 3(22), 470. https://doi.org/10.21105/joss.00470
