---
title: "The end of errors in ANOVA reporting"
layout: post
output:
  md_document:
    toc: yes
    variant: markdown_github
  html_document:
    df_print: paged
    toc: yes
author: "Dominique Makowski"
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---


```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
```



# Fit an anova

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
library(tidyverse)

simulate_logistic_data <- function(n=1000, noise=0){
   x <- rnorm(n)  # some continuous variables 
   z <- x  # linear combination with a bias
   pr <- 1/(1+exp(-z))  # pass through an inv-logit function
   if(noise != 0){
     pr <- pr + rnorm(n, 0, noise*sd(pr))
     pr[pr > 1] <- 1
     pr[pr < 0] <- 0
   }
   y <- rbinom(n, 1, pr)  # bernoulli response variable
   df <- data.frame(x=x, y=y)
   return(df)
}


data <- data.frame()
for(n in seq(10, 100)){
  print(n)
  
  for(i in seq(10)){
      df <- simulate_logistic_data(n, noise=0)
      
      fit = tryCatch({
        glm(y~x,data=df, family="binomial")
      }, warning = function(w) {
        0
      }, error = function(e) {
        0
      }, message = function(m) {
        0
      })
  
    if(is.numeric(fit)){
      next
    }
      
    for(method in c("chen2010", "cohen1988")){
      rez <- analyze(fit, effsize_rules = method)
      rez <- rez$summary[2,] %>% 
        select(Coef, Coef.std, p, Effect_Size) %>% 
        mutate(Method = method,
               n = n,
               Noise = noise,
               i = paste(i, n, sep="_"))
      data <- rbind(data, rez)
    }
  }
}

data %>% 
  mutate(Coef = abs(Coef),
         Effect_Size = forcats::fct_rev(Effect_Size)) %>% 
  ggplot(aes(y=Coef, x=Effect_Size, fill=Method)) +
  geom_violin() +
  ylim(0, 4)
```












# Contribute

Of course, these reporting standards should change, depending on new expert recommandations or official guidelines. **The goal of this package is to flexibly adaptive to new changes and good practices evolution**. Therefore, if you have any advices, opinions or such, we encourage you to either let us know by opening an [issue](https://github.com/neuropsychology/psycho.R/issues), or even better, try to implement them yourself by [contributing](https://github.com/neuropsychology/psycho.R/blob/master/.github/CONTRIBUTING.md) to the code. 


# Credits

This package helped you? Don't forget to cite the various packages you used :)

You can cite `psycho` as follows:

- Makowski, (2018). *The psycho Package: An Efficient and Publishing-Oriented Workflow for Psychological Science*. Journal of Open Source Software, 3(22), 470. https://doi.org/10.21105/joss.00470


# Previous blogposts

- [APA Formatted Bayesian Correlation](https://neuropsychology.github.io/psycho.R/2018/06/11/bayesian_correlation.html)
- [Fancy Plot (with Posterior Samples) for Bayesian Regressions](https://neuropsychology.github.io/psycho.R/2018/06/03/plot_bayesian_model.html)
- [How Many Factors to Retain in Factor Analysis](https://neuropsychology.github.io/psycho.R/2018/05/24/n_factors.html)
- [Beautiful and Powerful Correlation Tables](https://neuropsychology.github.io/psycho.R/2018/05/20/correlation.html)
- [Format and Interpret Linear Mixed Models](https://neuropsychology.github.io/psycho.R/2018/05/10/interpret_mixed_models.html)
- [How to do Repeated Measures ANOVAs](https://neuropsychology.github.io/psycho.R/2018/05/01/repeated_measure_anovas.html)
- [Standardize (Z-score) a dataframe](https://neuropsychology.github.io/psycho.R/2018/03/29/standardize.html)
- [Compute Signal Detection Theory Indices](https://neuropsychology.github.io/psycho.R/2018/03/29/SDT.html)
