---
title: "Modelling Curves: Fitting Polynomials Models"
layout: post
output:
  md_document:
    toc: yes
    variant: markdown_github
  html_document:
    df_print: paged
    toc: yes
author: "Dominique Makowski"
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---


```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
```


# Observed Data

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
library(tidyverse)
library(psycho)


df <- psycho::affective 
```


```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
plot <- df %>% 
  ggplot(aes(x=Adjusting, y=Life_Satisfaction)) +
  geom_count() +
  geom_smooth(method="loess", color="black", fill="black", alpha=0.1) +
  theme_classic()
plot
```



# Linear Fit

## Model

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='hide', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
fit <- lm(Life_Satisfaction ~ Adjusting, data=df)

analyze(fit) 
```

## Visualize

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
data_linear <- df %>% 
  select(Adjusting) %>% 
  refdata(length.out=100) %>% 
  get_predicted(fit, .)

plot <- plot +
  geom_ribbon(data=data_linear, 
              aes(y=Life_Satisfaction_Predicted, ymin=Life_Satisfaction_CI_2.5, ymax=Life_Satisfaction_CI_97.5), 
              fill="red", 
              alpha=0.1) +
  geom_line(data=data_linear, 
            aes(y=Life_Satisfaction_Predicted), 
            color="red")
plot
```

# Polynomial fit

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='hide', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
fit <- lm(Life_Satisfaction ~ poly(Adjusting, 2), data=df)

analyze(fit) 
```

## Visualize

To add the predictions of this model, we use the exact same code as for the linear model.

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
data_poly <- df %>% 
  select(Adjusting) %>% 
  refdata(length.out=100) %>% 
  get_predicted(fit, .)

plot <- plot +
  geom_ribbon(data=data_poly, 
              aes(y=Life_Satisfaction_Predicted, ymin=Life_Satisfaction_CI_2.5, ymax=Life_Satisfaction_CI_97.5), 
              fill="blue", 
              alpha=0.1) +
  geom_line(data=data_poly, 
            aes(y=Life_Satisfaction_Predicted), 
            color="blue")
plot
```


# Raw and Orthogonal Polynomials: Interpretation


```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='hide', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
data <- data.frame(x=seq(from=1, to=5, length.out=100),
                   y=4 - 0.6*x + 0.1*x^2 + 0.25*rnorm(100))

ggplot(data, aes(x, y)) + geom_point() + 
       geom_smooth(method = "lm", formula = y ~ poly(x, 2))

summary(lm(y ~ poly(x, 2, raw=TRUE)))
summary(lm(y ~ poly(x, 2)))
```

- Use raw polynomials if you're interested in the coefficient and/or the regression equation.
- Use orthogonal if you're interested in the general shape of the relationship.

# Contribute

**psycho is a young package in need of affection**. Therefore, if you have any advices, opinions or such, we encourage you to either let us know by opening an [issue](https://github.com/neuropsychology/psycho.R/issues), or by [contributing](https://github.com/neuropsychology/psycho.R/blob/master/.github/CONTRIBUTING.md) to the code. 


# Credits

This package helped you? Don't forget to cite the various packages you used :)

You can cite `psycho` as follows:

- Makowski, (2018). *The psycho Package: An Efficient and Publishing-Oriented Workflow for Psychological Science*. Journal of Open Source Software, 3(22), 470. https://doi.org/10.21105/joss.00470


# Previous blogposts

- [APA Formatted Bayesian Correlation](https://neuropsychology.github.io/psycho.R/2018/06/11/bayesian_correlation.html)
- [Fancy Plot (with Posterior Samples) for Bayesian Regressions](https://neuropsychology.github.io/psycho.R/2018/06/03/plot_bayesian_model.html)
- [How Many Factors to Retain in Factor Analysis](https://neuropsychology.github.io/psycho.R/2018/05/24/n_factors.html)
- [Beautiful and Powerful Correlation Tables](https://neuropsychology.github.io/psycho.R/2018/05/20/correlation.html)
- [Format and Interpret Linear Mixed Models](https://neuropsychology.github.io/psycho.R/2018/05/10/interpret_mixed_models.html)
- [How to do Repeated Measures ANOVAs](https://neuropsychology.github.io/psycho.R/2018/05/01/repeated_measure_anovas.html)
- [Standardize (Z-score) a dataframe](https://neuropsychology.github.io/psycho.R/2018/03/29/standardize.html)
- [Compute Signal Detection Theory Indices](https://neuropsychology.github.io/psycho.R/2018/03/29/SDT.html)
