---
title: "Specifying Random Effects in Psychology"
layout: post
output:
  md_document:
    toc: yes
    variant: markdown_github
  html_document:
    df_print: paged
    toc: yes
author: "Dominique Makowski"
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---


```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
```




# The data

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='hide', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
library(tidyverse)
library(psycho)

df <- psycho::emotion %>% 
  mutate(Trial_Order = as.factor(Trial_Order)) %>% 
  standardize()
```



# Fit Models
```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
library(lme4)
library(optimx)

fit_model <- function(random){
  fit <- lme4::glmer(paste("Recall ~ Emotion_Condition / Subjective_Arousal + ", random),
              data=df, 
              family = "binomial", 
              control = glmerControl(optimizer = "optimx", 
                                     calc.derivs = FALSE,
                                     optCtrl = list(method = "nlminb")))
  return(fit)
}



fit1 <- fit_model(random="(1|Participant_ID)") 
analyze(fit1, CI=NULL)

fit1 <- fit_model(random="(1|Participant_ID) + (1|Item_Name)") 
analyze(fit1, CI=NULL)

fit1 <- fit_model(random="(1|Participant_ID) + (1|Item_Name) + (1+Emotion_Condition|Participant_ID)") 
analyze(fit1, CI=NULL)

```


# Conclusion



# Contribute

Of course, these reporting standards should change, depending on new expert recommandations or official guidelines. **The goal of this package is to flexibly adaptive to new changes and good practices evolution**. Therefore, if you have any advices, opinions or such, we encourage you to either let us know by opening an [issue](https://github.com/neuropsychology/psycho.R/issues), or even better, try to implement them yourself by [contributing](https://github.com/neuropsychology/psycho.R/blob/master/.github/CONTRIBUTING.md) to the code. 


# Credits

This package helped you? Don't forget to cite the various packages you used :)

You can cite `psycho` as follows:

- Makowski, (2018). *The psycho Package: An Efficient and Publishing-Oriented Workflow for Psychological Science*. Journal of Open Source Software, 3(22), 470. https://doi.org/10.21105/joss.00470


# Previous blogposts

- [APA Formatted Bayesian Correlation](https://neuropsychology.github.io/psycho.R/2018/06/11/bayesian_correlation.html)
- [Fancy Plot (with Posterior Samples) for Bayesian Regressions](https://neuropsychology.github.io/psycho.R/2018/06/03/plot_bayesian_model.html)
- [How Many Factors to Retain in Factor Analysis](https://neuropsychology.github.io/psycho.R/2018/05/24/n_factors.html)
- [Beautiful and Powerful Correlation Tables](https://neuropsychology.github.io/psycho.R/2018/05/20/correlation.html)
- [Format and Interpret Linear Mixed Models](https://neuropsychology.github.io/psycho.R/2018/05/10/interpret_mixed_models.html)
- [How to do Repeated Measures ANOVAs](https://neuropsychology.github.io/psycho.R/2018/05/01/repeated_measure_anovas.html)
- [Standardize (Z-score) a dataframe](https://neuropsychology.github.io/psycho.R/2018/03/29/standardize.html)
- [Compute Signal Detection Theory Indices](https://neuropsychology.github.io/psycho.R/2018/03/29/SDT.html)
