---
title: "APA Formatted Correlation with effect sizes"
layout: post
output:
  md_document:
    toc: yes
    variant: markdown_github
  html_document:
    df_print: paged
    toc: yes
author: "Dominique Makowski"
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---


```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
```



To make sense of their data and effects, psychologists often standardize their variables. However, in repeated-measures designs, there are three ways of standardizing data:
- **variable-wise**: The most common method. A simple scaling and reducing of each variable.
- **participant-wise**: Variables are standardized "within" each participant.


In the following, we will see how to perform those methods and check for differences.

# The data

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='hide', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
library(tidyverse)
library(psycho)

df <- psycho::emotion %>% 
  filter(Emotion_Condition == "Negative")

df %>% 
  group_by(Participant_ID) %>% 
  summarise(n_Trials = n(),
            Subjective_Arousal_Mean = format_digit(mean(Subjective_Arousal, na.rm=TRUE)),
            Subjective_Arousal_SD = format_digit(sd(Subjective_Arousal, na.rm=TRUE)),
            Autobiographical_Link_Mean = format_digit(mean(Autobiographical_Link, na.rm=TRUE)),
            Autobiographical_Link_SD = format_digit(sd(Autobiographical_Link, na.rm=TRUE))) %>% 
  mutate(Subjective_Arousal = paste(Subjective_Arousal_Mean, Subjective_Arousal_SD, sep=" +- "),
         Autobiographical_Link = paste(Autobiographical_Link_Mean, Autobiographical_Link_SD, sep=" +- ")) %>% 
  select(-ends_with("SD"), -ends_with("Mean"))
```
```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
df %>% 
  group_by(Participant_ID) %>% 
  summarise(n_Trials = n(),
            Subjective_Arousal_Mean = format_digit(mean(Subjective_Arousal, na.rm=TRUE)),
            Subjective_Arousal_SD = format_digit(sd(Subjective_Arousal, na.rm=TRUE)),
            Autobiographical_Link_Mean = format_digit(mean(Autobiographical_Link, na.rm=TRUE)),
            Autobiographical_Link_SD = format_digit(sd(Autobiographical_Link, na.rm=TRUE))) %>% 
  mutate(Subjective_Arousal = paste(Subjective_Arousal_Mean, Subjective_Arousal_SD, sep=" +- "),
         Autobiographical_Link = paste(Autobiographical_Link_Mean, Autobiographical_Link_SD, sep=" +- ")) %>% 
  select(-ends_with("SD"), -ends_with("Mean")) %>%  
  knitr::kable() %>% 
  head(10)
```



# Standardize
```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
dfZ_VarWise <- df %>% 
  standardize()

dfZ_ParWise <- df %>% 
  group_by(Participant_ID) %>% 
  standardize()

dfZ_Full <- df %>% 
  group_by(Participant_ID) %>% 
  standardize() %>% 
  ungroup() %>% 
  standardize()
```



```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
print_summary <- function(data, var){
  paste(deparse(substitute(data)), ":", 
        format_digit(mean(data[[var]])),
        "+-",
        format_digit(sd(data[[var]])),
        "[", format_digit(min(data[[var]])),
        ",", format_digit(max(data[[var]])),
        "]")
}
print_summary(dfZ_VarWise, "Subjective_Arousal")
print_summary(dfZ_ParWise, "Subjective_Arousal")
print_summary(dfZ_Full, "Subjective_Arousal")
```



```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
print_participants <- function(data){
  data %>% 
    group_by(Participant_ID) %>% 
    summarise(Mean = mean(Subjective_Arousal), 
              SD = sd(Subjective_Arousal)) %>% 
    mutate_if(is.numeric, round, 2) %>% 
    head() 
    
}
print_participants(dfZ_VarWise)
print_participants(dfZ_ParWise)
print_participants(dfZ_Full)
```





```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
data.frame(VarWise = dfZ_VarWise$Subjective_Arousal,
           ParWise = dfZ_ParWise$Subjective_Arousal,
           Full = dfZ_Full$Subjective_Arousal) %>% 
  gather(Method, Variable) %>% 
  ggplot(aes(x=Variable, fill=Method)) +
  geom_density(alpha=0.2) +
  theme_minimal()
```

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
data.frame(Original = df$Subjective_Arousal,
           VarWise = dfZ_VarWise$Subjective_Arousal,
           ParWise = dfZ_ParWise$Subjective_Arousal,
           Full = dfZ_Full$Subjective_Arousal) %>% 
  ggplot(aes(x=Original)) +
  geom_point(aes(y=VarWise)) +
  geom_point(aes(y=ParWise), colour="red") +
  geom_point(aes(y=Full), colour="blue") +
  theme_minimal()
```


# Test

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
print_model <- function(data){
  lmerTest::lmer(Subjective_Arousal ~ Autobiographical_Link + (1|Participant_ID), data=data) %>% 
    analyze() %>% 
    summary() %>% 
    filter(Variable == "Autobiographical_Link") %>% 
    mutate(Type = deparse(substitute(data))) %>% 
    select(Type, Coef.std, p)
}

rbind(print_model(df), 
      print_model(dfZ_VarWise),
      print_model(dfZ_ParWise),
      print_model(dfZ_Full))
```

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
print_model <- function(data){
  lmerTest::lmer(Subjective_Arousal ~ Subjective_Valence + (1|Participant_ID), data=data) %>% 
    analyze() %>% 
    summary() %>% 
    filter(Variable == "Subjective_Valence") %>% 
    mutate(Type = deparse(substitute(data))) %>% 
    select(Type, Coef.std, p)
}

rbind(print_model(df), 
      print_model(dfZ_VarWise),
      print_model(dfZ_ParWise),
      print_model(dfZ_Full))
```

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
print_model <- function(data){
  lmerTest::lmer(Subjective_Valence ~ Autobiographical_Link + (1|Participant_ID), data=data) %>% 
    analyze() %>% 
    summary() %>% 
    filter(Variable == "Subjective_Valence") %>% 
    mutate(Type = deparse(substitute(data))) %>% 
    select(Type, Coef.std, p)
}

rbind(print_model(df), 
      print_model(dfZ_VarWise),
      print_model(dfZ_ParWise),
      print_model(dfZ_Full))
```


```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
print_model <- function(data){
  lme4::glmer(Recall ~ Subjective_Arousal + (1|Participant_ID), data=data, family="binomial") %>% 
    analyze() %>% 
    summary() %>% 
    filter(Variable == "Subjective_Arousal") %>% 
    mutate(Type = deparse(substitute(data))) %>% 
    select(Type, Coef.std, p)
}

rbind(print_model(df), 
      print_model(dfZ_VarWise),
      print_model(dfZ_ParWise),
      print_model(dfZ_Full))
```

# Contribute

Of course, these reporting standards should change, depending on new expert recommandations or official guidelines. **The goal of this package is to flexibly adaptive to new changes and good practices evolution**. Therefore, if you have any advices, opinions or such, we encourage you to either let us know by opening an [issue](https://github.com/neuropsychology/psycho.R/issues), or even better, try to implement them yourself by [contributing](https://github.com/neuropsychology/psycho.R/blob/master/.github/CONTRIBUTING.md) to the code. 


# Credits

This package helped you? Don't forget to cite the various packages you used :)

You can cite `psycho` as follows:

- Makowski, (2018). *The psycho Package: An Efficient and Publishing-Oriented Workflow for Psychological Science*. Journal of Open Source Software, 3(22), 470. https://doi.org/10.21105/joss.00470


# Previous blogposts

- [APA Formatted Bayesian Correlation](https://neuropsychology.github.io/psycho.R/2018/06/11/bayesian_correlation.html)
- [Fancy Plot (with Posterior Samples) for Bayesian Regressions](https://neuropsychology.github.io/psycho.R/2018/06/03/plot_bayesian_model.html)
- [How Many Factors to Retain in Factor Analysis](https://neuropsychology.github.io/psycho.R/2018/05/24/n_factors.html)
- [Beautiful and Powerful Correlation Tables](https://neuropsychology.github.io/psycho.R/2018/05/20/correlation.html)
- [Format and Interpret Linear Mixed Models](https://neuropsychology.github.io/psycho.R/2018/05/10/interpret_mixed_models.html)
- [How to do Repeated Measures ANOVAs](https://neuropsychology.github.io/psycho.R/2018/05/01/repeated_measure_anovas.html)
- [Standardize (Z-score) a dataframe](https://neuropsychology.github.io/psycho.R/2018/03/29/standardize.html)
- [Compute Signal Detection Theory Indices](https://neuropsychology.github.io/psycho.R/2018/03/29/SDT.html)
