---
title: "Bayesian Analyses in Psychology"
output: 
  rmarkdown::html_vignette:
    toc: true
author: 
- Dominique Makowski
date: "`r Sys.Date()`"
tags: [r, psychology, neuroscience]
abstract: |
  Why use frequentist methods when you can use, in an even simpler way, the Bayesian framework? Throughout this tutorial, we will explore many of the analyses you might want to do with your data.
vignette: >
  %\VignetteIndexEntry{BayesianPsychology}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{tidyverse}
  \usepackage[utf8]{inputenc}
---



------

```{r, echo=F, message=FALSE, warning=FALSE}
library(knitr)
```


## Why use the Bayesian Framework?

In short, because it's:

- Better
- Simpler
- Superior
- Preferable
- Above
- More appropriate
- More desirable
- More useful
- More valuable

##### **From Makowski et al. (*under review*):**

> Reasons to prefer this approach is better accuracy in noisy data, the possibility of introducing prior knowledge into the analysis and, critically, results intuitiveness and their straightforward interpretation. Indeed, in the frequentist view, the effects are fixed (but unknown) and data are random, while the Bayesian inference calculates the probability of different effect values (called the **"posterior" distribution**) given the observed data. Bayesian’s uncertainty can be summarized by giving a range of values on the posterior distribution that includes 95% of the probability (the 95% *Credible Interval*). To illustrate the difference, the Bayesian framework allows to say "*given the observed data, the effect has 95% probability of falling within this range*", while the Frequentist less straightforward equivalent would be "*there is a 95% probability that when computing a confidence interval from data of this sort, the effect falls within this range*".

## What is the Bayesian Framework?

Once we agreed that the Bayesian framework is the right way to go, you might wonder what is the Bayesian framework. **What's all the fuss about?**

Omitting the maths behind it, let's just say that:

- The frequentist guy tries to estimate "the real effect". The "real" value of the correlation between X and Y. It returns a "point-estimate" (*i.e.*, a single value) of the "real" correlation (*e.g.*, r = 0.42), considering that the data is sampled at random from a "parent", usually normal distribution of data.
- **The Bayesian master assumes not such thing**. The data are what they are. Based on this observed data (and eventually from its expectations), the Bayesian sampling algorithm will return a probability distribution of the effect that is compatible with the observed data. For the correlation between X and Y, it will return distribution that says "the most probable effect is 0.42, but this data is also compatible with a correlation of 0.12 or 0.74".
- To characterize our effects, **no need of p values** or other mysterious indices. We simply describe the posterior distribution (*i.e.*, the distribution of the effect). We can present the median (better than the mean, as it actually means that the effect has 50% of chance of being higher and 50% of chance of being lower), the MAD (a median-based, robust equivalent of SD) and other stuff such as the 90% HDI, called here *credible interval*.

**Let's practice!**

## The affective Dataset

Let's start by taking a look at the dataset included within the `psycho` package.

```{r, echo=T, message=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(rstanarm)
library(emmeans)
library(coda)
library(psycho)

df <- psycho::affective
summary(df)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(df)
```

The data includes **5 continuous variables** (age, life satisfaction and 3 affective styles) and **3 factors** (sex, salary and season of birth).

## Simple Regression (*Correlation*)

Let's start with something simple : a **correlation**. To simplify, a (Pearson's) correlation is pretty much nothing more than a simple linear regression (with standardized variables). Let's see if there's a linear relationship between **Life Satisfaction** and the tendency of **Tolerating** using a Bayesian linear regression model.

### Model Exploration

```{r, message=FALSE, results="hide"}
# Let's fit our model
fit <- rstanarm::stan_glm(Life_Satisfaction ~ Tolerating, data=df)
```




Let's check the results: 
```{r, message=FALSE, results="hide"}
# Format the results using analyze()
results <- psycho::analyze(fit)

# We can extract a formatted summary table
summary(results, round = 2)
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(summary(results, round = 2))
```

And print it textually:
```{r echo=T, message=FALSE, warning=FALSE}
print(results)
```

### Interpretation


> Full Bayesian mixed linear models are fitted using the rstanarm R wrapper for the stan probabilistic language (Gabry & Goodrich, 2016). Bayesian inference was done using Markov Chain Monte Carlo (MCMC) sampling. The prior distributions of all effects were set as weakly informative, meaning that we did not expect effects different from null in any particular direction. For each model and each coefficient, we will present several characteristics of the posterior distribution, such as its median (a robust estimate comparable to the beta from frequentist linear models), MAD (median absolute deviation, a robust equivalent of standard deviation) and the 95% credible interval. Instead of the p value as an index of effect existence, we also computed the maximum probability of effect (MPE), *i.e.*, the maximum probability that the effect is different from 0 in the median’s direction. An effect can be considered as inconsistant if its MPE is lower than 90% (however, **caution not to fall in the *p* value problem**).


The current model explains about 2.3% of life satisfaction variance. Within this model, a positive linear relationship between life satisfaction and tolerating exists with high probability (Median = 0.19, MAD = 0.035, 95% CI [0.12, 0.26], MPE = 100%).

### Model Visualization

To visualize the model, we have to extract the "reference grid" (a theorethical dataframe with balanced data), using the `emmeans` package.

```{r echo=T, message=FALSE, warning=FALSE}
# We enter the values of tolerating that we want in our reference grid, in this case a sequence of length=10 with minimum and maximum similar to the actual data.
ref_grid <- emmeans::ref_grid(fit, at = list(
  Tolerating = seq(min(df$Tolerating),
                   max(df$Tolerating),
                   length.out = 10)))
                              
predicted <- psycho::get_predicted(fit, refgrid=ref_grid)

```
```{r echo=T, message=FALSE, warning=FALSE, results='hide'}
predicted
```




```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(predicted)
```

Our refgrid is made of equally spaced (balanced) predictor values. It also include the predicted outcome's predictions, as well as 90% credible intervals. Now, we can plot it as follows:

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA}
ggplot(predicted, aes(x=Tolerating, y=pred_Life_Satisfaction)) +
  geom_line() +
  geom_ribbon(aes(ymin=`pred_Life_Satisfaction_5%`, 
                  ymax=`pred_Life_Satisfaction_95%`), 
              alpha=0.1)

```


## Regression with Categorical Predictor (*ANOVA*)

When the predictor is categorical, simplifying the model is called running an ANOVA. Let's do it by answering the following question: does the level of **life satisfaction** depend on the salary? 

### Model Exploration

```{r, message=FALSE, results="hide"}
# Let's fit our model
fit <- rstanarm::stan_glm(Life_Satisfaction ~ Salary, data=df)
```
Let's check the results: 
```{r, message=FALSE, results="hide"}
# Format the results using analyze()
results <- psycho::analyze(fit)

# We can extract a formatted summary table
summary(results, round = 2)
```


### Post-hoc / Contrasts / Comparisons

What interest us is the pairwise comparison between the groups. The `get_contrasts` function computes the estimated marginal means (least-squares means), *i.e.*, the means of each group estimated by the model, as well as the contrasts.

```{r, message=FALSE, results="hide"}
contrasts <- psycho::get_contrasts(fit, "Salary")
```

We can see the estimated means like that:
```{r echo=T, message=FALSE, warning=FALSE, results='hide'}
contrasts$means
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(contrasts$means)
```


And the contrasts comparisons like that:
```{r echo=T, message=FALSE, warning=FALSE, results='hide'}
contrasts$contrasts
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(contrasts$contrasts)
```

As we can see, the only probable difference (MPE > 90%) is between **Salary <1000** and **Salary 2000+**.

### Model Visualization

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA}
ggplot(contrasts$means, aes(x=Contrast, y=Median, group=1)) +
  geom_line() +
  geom_pointrange(aes(ymin=CI_lower, ymax=CI_higher)) +
  ylab("Life Satisfaction") +
  xlab("Salary")
```


## Logistic Regressions

Let's see if we can **predict the sex** with the tendency to flexibly *adjust* our emotional reactions. As the Sex is a binary factor (with two modalities), we have to fit a logistic model.

### Model Exploration

```{r, message=FALSE, results="hide"}
# Let's fit our model
fit <- rstanarm::stan_glm(Sex ~ Adjusting, data=df, family = binomial(link="logit"))
```



First, let's check our model: 
```{r, message=FALSE, results="hide"}
# Format the results using analyze()
results <- psycho::analyze(fit)

# We can extract a formatted summary table
summary(results, round = 2)
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(summary(results, round = 2))
```

It appears that the link between adjusting and the sex is highly probable (MPE > 90%). But in what direction? To know that, we have to find out what is the intercept (the reference level).


```{r echo=FALSE, message=FALSE, warning=FALSE}
levels(df$Sex)
```
As **female** is the first level, it means that it is the intercept. Based on our model, an increase of 1 on the scale of **adjusting** will increase the probability (expressed in log odds ratios) of being a **male**.

### Model Visualization

To visualize this type of model, we have to derive a reference grid.

```{r echo=T, message=FALSE, warning=FALSE}
ref_grid <- emmeans::ref_grid(fit, at = list(
  Adjusting = seq(min(df$Adjusting),
                   max(df$Adjusting),
                   length.out = 100)))

predicted <- psycho::get_predicted(fit, refgrid=ref_grid)
```

However, as the predicted values is expressed in log odds, it's better to convert it to probability, which make more sense to us Humans.

```{r echo=FALSE, message=FALSE, warning=FALSE}
predicted$pred_Sex_proba <- psycho::odds_to_probs(predicted$pred_Sex)
predicted$`pred_Sex_5%_proba` <- psycho::odds_to_probs(predicted$`pred_Sex_5%`)
predicted$`pred_Sex_95%_proba` <- psycho::odds_to_probs(predicted$`pred_Sex_95%`)
```

Now we can plot it.

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA}
ggplot(predicted, aes(x=Adjusting, y=pred_Sex_proba)) +
  geom_line() +
  geom_ribbon(aes(ymin=`pred_Sex_5%_proba`, 
                  ymax=`pred_Sex_95%_proba`), 
              alpha=0.1) +
  ylab("Probability of being a male")
```

We can nicely see the non-linear relationship between adjusting and the probability of being a male.

## Multiple Regressions and MANOVAs / ANCOVAs

Let's create models a bit more complex, mixing factors with numeric predictors, to see if the **life satisfaction** is related to the tendency to suppress, **conceal** the emotional reactions, and does this relationship depends on the **sex**.

### Model Exploration

```{r, message=FALSE, results="hide"}
# Let's fit our model
fit <- rstanarm::stan_glm(Life_Satisfaction ~ Concealing * Sex, data=df)
```



Let's check our model: 
```{r, message=FALSE, results="hide"}
# Format the results using analyze()
results <- psycho::analyze(fit)

# We can extract a formatted summary table
summary(results, round = 2)
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(summary(results, round = 2))
```

Again, it is important to notice that the intercept (the baseline) corresponds here to **Concealing = 0** and **Sex = F**. As we can see next, there is, with high probability, a negative linear relationship between concealing (*for females only*) and life satisfaction. Also, at the (theorethical) intercept (when concealing = 0), the males have a lower life satisfaction. Finally, the interaction is also probable. This means that when the participant is a male, the relationship between concealing and life satisfaction is significantly different (increased by 0.17. In other words, we could say that the relationship is of -0.10+0.17=0.07 in men).

### Model Visualization

How to represent this type of models? Again, we have to generate a reference grid.

```{r echo=T, message=FALSE, warning=FALSE}
ref_grid <- emmeans::ref_grid(fit, at = list(
  Concealing = seq(min(df$Concealing),
                   max(df$Concealing),
                   length.out = 5)))
predicted <- psycho::get_predicted(fit, refgrid=ref_grid)
```
```{r echo=T, message=FALSE, warning=FALSE, results='hide'}
predicted
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(predicted)
```

As we can see, the reference grid is balanced in terms of factors and numeric predictors. Now, to plot this becomes very easy !


```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA}
ggplot(predicted, aes(x=Concealing, y=pred_Life_Satisfaction, colour=Sex, fill=Sex)) +
  geom_line() +
  geom_ribbon(aes(ymin=`pred_Life_Satisfaction_5%`, 
                  ymax=`pred_Life_Satisfaction_95%`), 
              alpha=0.1)
```

First, we can see that the error for the males is larger, due to less observations. Second, we can see that the plot is a bit edgy, and we can smooth it by generating a more detailed reference grid.

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA}
ref_grid <- emmeans::ref_grid(fit, at = list(
  Concealing = seq(min(df$Concealing),
                   max(df$Concealing),
                   length.out = 100)))  # Increase this number to get more observations
predicted <- psycho::get_predicted(fit, refgrid=ref_grid)


ggplot(predicted, aes(x=Concealing, y=pred_Life_Satisfaction)) +
  geom_line(aes(colour=Sex), size=2) +
  geom_ribbon(aes(ymin=`pred_Life_Satisfaction_5%`, 
                  ymax=`pred_Life_Satisfaction_95%`,
                  fill=Sex), 
              alpha=0.2) +
  ylab("Life Satisfaction")
```



## Mixed Models


### Why use mixed-models?

- **From Makowski et al. (*under review*):**

> The Mixed modelling framework allows estimated effects to vary by group at lower levels while estimating population-level effects through the specification of fixed (explanatory variables) and random (variance components) effects. Outperforming traditional procedures such as repeated measures ANOVA (Kristensen & Hansen, 2004), these models are particularly suited to cases in which experimental stimuli are heterogeneous (e.g., images) as the item-related variance, in addition to the variance induced by participants, can be accounted for (Baayen, Davidson, & Bates, 2008; Magezi, 2015). Moreover, mixed models can handle unbalanced data, nested designs, crossed random effects and missing data.

As for how to run this type of analyses, it is quite easy. Indeed, all what has been said previously remains the same for mixed models. Except that there are random effects (specified by putting `+ (1|random_term)` in the formula). For example, we might want to consider the **salary** as a random effect (to "**adjust**" (*so to speak*) for the fact that the data is structured in two groups). Let's explore the relationship between the tendency to **conceal** emotions and **age** (*adjusted* for **salary**).

### Model Exploration

```{r, message=FALSE, results="hide", warning=FALSE}
# Let's fit our model (it takes more time)
fit <- rstanarm::stan_lmer(Concealing ~ Age + (1|Salary), data=df)
```

Let's check our model: 
```{r, message=FALSE, results="hide"}
# Format the results using analyze()
results <- psycho::analyze(fit)

# We can extract a formatted summary table
summary(results, round = 2)
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(summary(results, round = 2))
```

As we can see, the linear relationship is probable.

### Model Visualization


```{r echo=T, message=FALSE, warning=FALSE}
ref_grid <- emmeans::ref_grid(fit, at = list(
  Age = seq(min(df$Age),
            max(df$Age),
            length.out = 10)))
# We name the predicted dataframe by adding _linear to keep it for further comparison (see next part)
predicted_linear <- psycho::get_predicted(fit, refgrid=ref_grid)

```
```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA}
ggplot(predicted_linear, aes(x=Age, y=pred_Concealing)) +
  geom_line() +
  geom_ribbon(aes(ymin=`pred_Concealing_5%`, 
                  ymax=`pred_Concealing_95%`), 
              alpha=0.1)
```

## Polynomial Transformations

Relationships in the real world are often non-linear. For example, based on the previous relationship between **concealing** and **age**, we could try modelling a polynomial (second order) transformation to the predictor.

### Model Exploration

```{r, message=FALSE, results="hide", warning=FALSE}
# Let's fit our model (it takes more time)
fit <- rstanarm::stan_lmer(Concealing ~ poly(Age, 2, raw=TRUE) + (1|Salary), data=df)
```

Let's check our model: 
```{r, message=FALSE, results="hide"}
# Format the results using analyze()
results <- psycho::analyze(fit)

# We can extract a formatted summary table
summary(results, round = 2)
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(summary(results, round = 2))
```

As we can see, both the linear relationship and the second order curvature are highly probable. However, when setting `raw=TRUE` in the formula, the coefficients become unintepretable. So let's visualize them.

### Model Visualization

The model visualization routine is similar to the previous ones.

```{r echo=T, message=FALSE, warning=FALSE}
ref_grid <- emmeans::ref_grid(fit, at = list(
  Age = seq(min(df$Age),
            max(df$Age),
            length.out = 100)))

predicted_poly <- psycho::get_predicted(fit, refgrid=ref_grid)
```
```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA}
ggplot(predicted_poly, aes(x=Age, y=pred_Concealing)) +
  geom_line() +
  geom_ribbon(aes(ymin=`pred_Concealing_5%`, 
                  ymax=`pred_Concealing_95%`), 
              alpha=0.1)
```

As we can see, adding the polynomial degree changes the relationship. Since the model is here very simple, we can add on the plot the actual points (however, they do not take into account the random effects and such), as well as plot the two models. Also, let's make it "dynamic" using `plotly`.

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
p <- ggplot() +
  # Linear model
  geom_line(data=predicted_linear, 
            aes(x=Age, y=pred_Concealing),
            colour="blue",
            size=1) +
  geom_ribbon(data=predicted_linear, 
              aes(x=Age,
                  ymin=`pred_Concealing_5%`,
                  ymax=`pred_Concealing_95%`), 
              alpha=0.1,
              fill="blue") +
  # Polynormial Model
  geom_line(data=predicted_poly, 
            aes(x=Age, y=pred_Concealing),
            colour="red",
            size=1) +
  geom_ribbon(data=predicted_poly, 
              aes(x=Age,
                  ymin=`pred_Concealing_5%`, 
                  ymax=`pred_Concealing_95%`), 
              fill="red",
              alpha=0.1) +
  # Actual data
  geom_point(data=df, aes(x=Age, y=Concealing))

library(plotly) # To create interactive plots
ggplotly(p) # To transform a ggplot into an interactive plot
```


**It's good to take a few steps back and look at the bigger picture :)**

## Piors Specification

One of the interesting aspect of the Bayesian framework is the possibility of adding prior expectations about the effect, to help model fitting and increase accuracy in noisy data or small samples.


### Weakly informative priors

As you might have notice, we didn't specify any priors in the previous analyses. In fact, we let the algorithm define and set *weakly informative priors*, designed to provide moderate regularization and help stabilize computation, without biasing the effect direction. For example, a wealky informative prior, for a standardized predictor (with mean = 0 and SD = 1) could be a normal distribution with mean = 0 and SD = 1. This means that the effect of this predictor is expected to be equally probable in any direction (as the distribution is symmetric around 0), with probability being higher close to 0 and lower far from 0.

While this prior doesn't bias the direction of the Bayesian (MCMC) sampling, it suggests that having an effect of 100 (*i.e.*, located at 100 SD of the mean as our variables are standardized) is highly unprobable, and that an effect close to 0 is more probable.

To better play with priors, let's start by standardizing our dataframe.


```{r, message=FALSE, results="hide"}
# Standardize (scale and center) the numeric variables
dfZ <- psycho::standardize(df)
```

Then, we can explicitly specify a weakly informative prior for all effects of the model.

```{r, message=FALSE, results="hide"}
# Let's fit our model
fit <- rstanarm::stan_glm(Life_Satisfaction ~ Tolerating, 
                          data=dfZ,
                          prior=normal(location = 0, # Mean
                                       scale = 1)) # SD
```

Let's plot the prior (the expectation) against the posterior (the estimated effect) distribution.

```{r, message=FALSE, results="hide"}
results <- psycho::analyze(fit)

# Extract the posterior
posterior <- results$values$Tolerating$posterior

# Create a posterior with the prior and posterior distribution and plot them.
data.frame(posterior = posterior,
           prior = rnorm(length(posterior), 0, 1)) %>% 
  ggplot() +
  geom_density(aes(x=posterior), fill="lightblue", alpha=0.5) +
  geom_density(aes(x=prior), fill="blue", alpha=0.5) +
  scale_y_sqrt() # Change the Y axis so the plot is less ugly
```


This plot is rather ugly, because our posterior is very precise (due to the large sample) compared to the prior. 

### Informative priors

It is important to underline that setting informative priors (**if realistic**), does not overbias the analysis. In other words, is only "directs" the sampling. But the observed data still wins. **In other words, you can't change the results dramatically by tweaking the priors**. Of course, if you see someone using a prior with mean = 42 and SD = 0.0001, you should look at his results with caution...



## Credits

This package helped you? Don't forget to cite the various packages you used :)

You can cite `psycho` as follows:

- Makowski, (2018). **The psycho Package: an Efficient and Publishing-Oriented Workflow for Psychological Science**. Journal of Open Source Software, 3(22), 470. https://doi.org/10.21105/joss.00470
