---
title: "Generate posts"
layout: post
output:
  md_document:
    variant: markdown_github
author: "Dominique Makowski"
date: "`r Sys.Date()`"
tags:
- r
- psychology
- neuroscience
- neuropsychology
- standardize
- normalize
- z score
- scale and center
---



## Standardize / Normalize / Z-score / Scale

The `standardize()` function allows you to easily scale and center all numeric variables of a dataframe. It is similar to the base function `scale()`, but presents some advantages: it is tidyverse-friendly, data-type friendly (*i.e.*, does not transform it into a matrix) and can handle dataframes with categorical data.

```{r, out.width=8, eval = TRUE, fig.align='center', results='markup', comment=NA}
library(psycho)
library(tidyverse)

z_iris <- iris %>% 
  psycho::standardize() 

summary(z_iris)
```

But beware, standardization **does not** change (and "normalize") the distribution!

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center'}
z_iris %>% 
  dplyr::select(-Species) %>% 
  gather(Variable, Value) %>% 
  ggplot(aes(x=Value, fill=Variable)) +
      geom_density(alpha=0.5) +
      geom_vline(aes(xintercept=0)) +
      theme_bw() +
      scale_fill_brewer(palette="Spectral")
```
