---
title: "Generate posts"
layout: post
output:
  md_document:
    variant: markdown_github
author: "Dominique Makowski"
date: "`r Sys.Date()`"
tags:
- r
- psychology
- neuroscience
- neuropsychology
- Single case methods
- crawford-howell
- mellenbergh
---


## Single-case methods

### Crawford-Howell (1998) t-test for single-case vs. control group

Neuropsychologists (and neuropsychology students) often need to compare a single case to a small control group. However, the standard two-sample t-test does not work because the case is only one observation. Crawford and Garthwaite (2012) demonstrate that the Crawford-Howell (1998) t-test is a better approach (in terms of controlling Type I error rate) than other commonly-used alternatives.


```{r echo=TRUE, message=FALSE, warning=FALSE, results='markup'}
library(psycho)

case <- 82 # The IQ of a patient
controls <- c(86, 100, 112, 95, 121, 102) # The IQs of a control group

rez <- crawford.test(case, controls)
```


### Mellenbergh & van den Brink (1998) test for pre-post comparison

Clinicians willing to check if their intervention had an effect on a single participant might want to use the Mellenbergh & van den Brink (1998) test, comparing the difference between baseline and post-test to the standart deviation of a control group.

```{r echo=TRUE, message=FALSE, warning=FALSE, results='markup'}
library(psycho)

t0 <- 82 # The IQ of a patient at baseline
t1 <- 105 # The IQ of a patient after the new therapy
controls <- c(94, 100, 108, 95, 102, 94) # The IQs of a control group

rez <- mellenbergh.test(t0, t1, controls = controls)
```

If we do not have a control group, we can also directly enter the SD of the score. For IQ, the SD is of 15.

```{r echo=TRUE, message=FALSE, warning=FALSE, results='markup'}
rez <- mellenbergh.test(t0, t1, controls = 15)
```
