---
title: "Easy APA Formatted Bayesian Correlation"
layout: post
output:
  html_document:
    df_print: paged
    toc: yes
  md_document:
    toc: yes
    variant: markdown_github
author: "Dominique Makowski"
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---


```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
```

**The Bayesian framework is the right way to go for psychological science.** To facilitate its use for newcommers, we implemented the `bayes_cor.test` function in the [psycho package](https://github.com/neuropsychology/psycho.R), a **user-friendly wrapper** for the `correlationBF` function of the great [`BayesFactor`](https://richarddmorey.github.io/BayesFactor/) package by Richard D. Morey.

# Traditional Correlation

Let's first perform a traditional correlation.

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
# devtools::install_github("neuropsychology/psycho.R")  # Install the latest psycho version

# Load packages
library(tidyverse)
library(psycho)

# Import data
df <- psycho::affective

cor.test(df$Concealing, df$Tolerating)
```

There is a **significant** (*whatever that means*), yet **weak positive** correlation between Concealing and Tolerating affective styles.


# Bayesian APA formatted Correlation

And now, see how quickly we can do a Bayesian correlation: 

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='markup', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
bayes_cor.test(df$Concealing, df$Tolerating)
```

The results are roughly the same, but neatly dissociate the likelihood in favour or against the null hypothesis (using the [Bayes Factor](https://www.r-bloggers.com/what-does-a-bayes-factor-feel-like/)), from the "significance" (wether a portion of the *Credible Interval* covers 0), from the effect size (interpreted here with [Cohen's (1988) rules of thumb](https://github.com/neuropsychology/psycho.R/blob/master/R/interpret_r.R#L142)). Critically, **you can, now, simply copy/paste this output to your manuscript!** It includes and formats the Bayes Factor, the median (a good point-estimate, close to the *r* estimated in frequentist correlation), MAD (robust equivalent of SD for median) and *credible* interval (CI) of the posterior distribution, as well as effect size interpretation.

# Indices

We can have access to more indices with the `summary`:

```{r, fig.width=7, fig.height=4.5, eval = TRUE, results='hide', fig.align='center', comment=NA, message=FALSE, warning=FALSE}
results <- bayes_cor.test(df$Concealing, df$Tolerating)
summary(results)
```
```{r, fig.width=7, fig.height=4.5, echo=FALSE, eval = TRUE, fig.align='center', comment=NA, message=FALSE, warning=FALSE}
knitr::kable(summary(results), digits=2)
```

Those indices include the ROPE decision criterion (see [Kruschke, 2018](http://journals.sagepub.com/doi/abs/10.1177/2515245918771304)) as well as the Maximum Probability of Effect (MPE, the probability that an effect is negative or positive and different from 0).

# Posterior 

We can easily extract the posterior distribution to visualize the probability of possible effects.

```{r, fig.width=7, fig.height=4.5, eval = TRUE, fig.align='center', comment=NA, message=FALSE, warning=FALSE}
posterior <- results$values$posterior

plot(density(posterior))
```



# Credits

This package helped you? Don't forget to cite the various packages you used :)

You can cite `psycho` as follows:

- Makowski, (2018). *The psycho Package: an Efficient and Publishing-Oriented Workflow for Psychological Science*. Journal of Open Source Software, 3(22), 470. https://doi.org/10.21105/joss.00470
